Text embedding models reveal high resolution insights into conceptual knowledge from short multiple choice quizzes

### 2023-01-27 paper/main.tex

\begin{abstract} We develop a mathematical framework, based on natural language
processing models, for tracking and characterizing the acquisition of
conceptual knowledge. Our approach embeds each concept in a high dimensional
representation space, where nearby coordinates reflect similar or related
concepts. We test our approach using behavioral data collected from a group of
college students. In the experiment, we ask the participants to answer small
sets of multiple choice quiz questions interleaved between watching two course
videos from the Khan Academy platform. We applied our framework to the videos'
transcripts, and to text of the quiz questions, to quantify the content of each
moment of video and each quiz question. We used these embeddings, along with
participants' quiz responses, to track how the learners' knowledge changed
after watching each video. Our findings show how a small set of quiz questions
may be used to obtain rich and meaningful high resolution insights into what
each learner knows, and how their knowledge changes over time as they learn.

@@

We asked our participants to answer questions from several multiple choice
quizzes and watch two lecture videos from the \textit{Khan Academy} platform
(Fig.~\ref{fig:experiment}). One lecture video, entitled \textit{Four
Fundamental Forces}, was about the four fundamental forces in physics: gravity,
strong and weak interactions, and electromagnetism. The second lecture video,
entitled \textit{Birth of Stars}, provides an overview of our current
understanding of how stars form. We selected both lessons to be (a) accessible
to a broad audience, e.g., by minimizing prerequisite knowledge, (b) largely
independent of each other, e.g., so that the two videos focused on different
material and did not depend on each other, and (c) related to each other, e.g.,
so that both videos contained at least \textit{some} similar or overlapping
content. The two videos we selected are introductory (i.e., minimizing specific
prerequisite knowledge), and are about different primary concepts, but they
both also touch on ``physics'' and ``astronomy'' themes. We also wrote a set of
multiple choice quiz questions that would enable us to test participants'
knoweldge about each individual video and about related content not
specifically presented in either video (Tab.~\questions). Participants answered
questions randomly drawn from each content area (lecture 1, lecture 2, and
general physics knowledge) across each of three quizzes. Quiz 1 was intended to
assessed participants' knowledge before training; quiz 2 assessed knowledge
after watching the \textit{Four Fundamental Forces} video (i.e., lecture 1);
and quiz 3 assessed knowledge after watching the \textit{Birth of Stars} video
(i.e., lecture 2).

@@

Given that we trained the text embedding model using the video transcripts, we
wondered whether the questions that were (ostensibly, by design) ``about'' the
content of each lecture would ``match up'' correctly with the lectures. In
other words, we hoped that the text embeddings would capture something about
the deeper conceptual content of the lectures, beyond surface details such as
exact wording choices. If so, when we embed \textit{new} text outside of the
model's training set, we should see a correspondance between the embeddings of
the training data (i.e., snippets of text from the lectures' transcripts) and
other text that reflects related concepts (e.g., questions \textit{about} each
lecture). Further, although the content from any given moment from a lecture
might stray from the average content (across all timepoints), we hoped that
\textit{variability} in each topic's expression over timepoints within a
lecture would match up with the variability in topic expressions for questions
about that lecture. Intuitively, the variability in the expression of a given
topic relates to how much ``information''~\citep{Fish22} the lecture (or
questions) reflect about that topic. When we compared the variability in topic
weights across each lecture's timepoints with the variability in topic weights
across each question set, we found a strong correspondence
(Fig.~|ref{fig:topics}). The most variable topics from the \textit{Four
Fundamental Forces} lecture, and questions about that lecture, are 2, 3, 8, 12,
13, 14, and 15. The most variable topics from the \textbf{Birth of Stars}
lecture, and questions about that lecture, are 1, 4, 5, 6, and 7. This strong
overlap between the lectures and questions specifically about each lecture
indicates that the topic model captures some of the underlying conceptual
content.

@@

Although a single lecture may be organized around a single broad theme at a
coarse scale, at a finer scale each moment of a lecture typically covers a
narrower range of content. We wondered whether a text embedding model trained
on the lectures' transcripts might capture some of this finer scale content.
For example, if a particular question asks about the content from one small
part of a lecture, we wondered whether our text embedding model could be used
to automatically identify the ``matching'' moment(s) in the lecture. When we
correlated each question's topic vector with the topic vectors for each second
of the lectures, we found some evidence that each question is temporally
specific (Fig.~\ref{fig:question-correlations}). In particular, most questions'
topic vectors were maximally correlated with a well-defined (and relatively
narrow) range of timepoints from their corresponding lectures, and the
correlations fell off sharply outside of that range. We also examined the
best-matching intervals for each question qualitatively by comparing the text
of the question to the text of the most-correlated parts of the lectures.
Despite that the questions were excluded from the text embedding model's
training set, in general we found (through manual inspection) a close
correspondence between the conceptual content that each question covered and
the content covered by the best-matching moments of the lectures. Two
representative examples are shown at the bottom of
Fig.~\ref{fig:question-correlations}.

The ability to quantify how much each question is ``asking about'' the content
from each moment of the lectures could enable high-resolution insights into
participants' knowledge. Traditional approaches to estimating how much a
student ``knows'' about the content of a given lecture entail computing the
proportion of correctly answered questions. But if two students receive
identical scores on an exam, might our modeling framework help us to gain more
naunced insights into the \textit{specific} content that each student has
mastered (or failed to master)? For example, a student who misses three
questions that were all about the same concept (e.g., concept $A$) will have
gotten the same \textit{proportion} of questions correct as another student who
missed three questions about three \textit{different} concepts (e.g., $A$, $B$,
and $C$). But if we wanted to fill in the ``gaps'' in the two students'
understandings, we might do well to focus on concept $A$ for the first student,
but to also add in materials pertaining to concepts $B$ and $C$ for the second
student.

@@

If we are able to accurately estimate a participant's knowledge about the
content tested by a given question, the estimated knowledge should have some
predictive information about whether the participant is likely to answer the
question correctly or incorrectly. For each question in turn, for each
participant, we used Equation~\ref{eqn:prop} to estimate (using all
\textit{other} questions from the same quiz, from the same participant) the
participant's knowledge at the held-out question's embedding coordinate. For
each quiz, we aggregated these estimates into two distributions: one for the
estimated knowledge at the coordinates of each \textit{correctly} answered
question, and another for the estimated knowledge at the coordinates of each
\textit{incorrectly} answered question (Fig.~\ref{fig:predictions}). We then
used independent samples $t$-tests to compare the means of these distributions
of estimated knowledge.

@@

    \caption{\textbf{Mapping out the geometry of knowledge and learning.}
    \textbf{A. Average ``knowledge maps'' estimated using each quiz.} Each map
    displays a 2D projection of the estimated knowledge about the content
    reflected by \textit{all} regions of topic space (see \textit{Creating
    knowledge and learning map visualizations}). The topic trajectories of each
    lecture and the coordinates of each question are indicated by dotted lines
    and dots. Each map reflects an average across all participants. For
    individual participants' maps, see
    Figures~\individualKnowledgeMapsA,~\individualKnowledgeMapsB,
    and~\individualKnowledgeMapsC. \textbf{B. Average ``learning maps''
    estiamted between each successive pair of quizzes.} The learning maps are
    in the same general format as the knowledge maps in Panel A, but each
    coordinate in the learning maps indicates the \textit{difference} between
    the corresponding coordinates in the indicated \textit{pair} of knowledge
    maps---i.e., how much the estimated knowledge ``changed'' across the two
    quizzes. Each map reflects an average across all participants. For
    individual participants' maps, see
    Figures~\individualLearningMapsA~and~\individualLearningMapsB. \textbf{C.
    Word clouds for sampled points in topic space.} Each word cloud displays
    the relative weights of each word reflected by the blend of topics
    represented at the locations of the stars in the maps. The words' colors
    indicate how much each word is weighted on average across all timepoints'
    topic vectors in the \textit{Four Fundamental Forces} (blue) and
    \textit{Birth of Stars} (green) videos, respectively.}

@@

Because the 2D projection we used to construct the knowledge and learning maps
is (partially) invertable, we may gain additional insights into the estimates
by reconstructing the original high-dimensional topic vectors for any point(s)
in the maps we are interested in. For example, this could serve as a useful
tool for an instructor looking to better understand which content areas a
student (or a group of students) knows well (or poorly). As a demonstration, we
show the top-weighted words from the blends of topics reconstructed from three
example locations on the maps (Fig.~\ref{fig:knowledge-maps}C): one point near
the \textit{Four Fundamental Forces} embedding (yellow); a second point near
the \textit{Birth of Stars} embedding (orange), and a third point somewhere in
between the two lectures' embeddings (pink). As shown in the word clouds in the
Panel, the top-weighted words at the example coordinate near the \textit{Four
Fundamental Forces} embedding also tended to be weighted heavily by the topics
expressed in that lecture. Similarly, the top-weighted words at the example
coordinate near the \textit{Birth of Stars} embedding tended to be weighted
most heavily by the topics expressed in \textit{that} lecture. And the
top-weighted words at the example coordinate between the two lectures'
embeddings show a roughly even mix of words most strongly associated with each
lecture.

@@

The notion of empathy also related to ``theory of mind'' of other
individuals~\citep{GoldWinn12, KansEtal15, Melt11}. Considering others' unique
perspectives, prior experiences, knowledge, goals, etc., can help us to more
effectively interact and communicate~\citep{ShaoEtal18, StepBaer06, Ratk18}.
The knowledge and learning maps we estimate in our study
(Fig.~\ref{fig:knowledge-maps}) hint at one potential form that an automated
``empathetic'' teacher might take. We imagine that automated content delivery
systems that adapt lessons on the fly according to continually updated
estimates of what students know and how quickly they are learning different
conceptual content~\citep[e.g., building on ideas such as ][and
others]{AndeSkwa86, Half88, Kuma05, WolzEtal88}.

@@

Participants' ages ranged from 18 to 22 years (mean: 19.52 years;
standard deviation: 1.09 years). A total of 15 participants reported
their gender as male and 35 participants reported their gender as
female. A total of 49 participants reported their native language as
``English'' and 1 reported having another native language. A total of
47 participants reported their ethnicity as ``Not Hispanic or Latino''
and three reported their ethnicity as ``Hispanic or Latino.''
Participants reported their races as White (32 participants), Asian
(14 participants), Black or African American (5 participants),
American Indian or Alaska Native (1 participant), and Native Hawaiian or
Other Pacific Islander (1 participant). (Note that some participants
selected multiple recial categories.)

@@

Particpants reported their undergraduate major(s) as Social Sciences (28
participants), Natural sciences (16), Professional (e.g., pre-med or pre-law; 8
participants), Mathematics and engineering (7 participants), Humanities (4
participants), or Undecided (3 participants). Note that some participants
selected multiple categories for their undergraduate major. We also asked
participants about the courses they had taken. In total, 46 participants
reported having taken at least one Khan academy course in the past or being
familiar with the Khan academy, and 4 reported not having taken any Khan
academy courses. Of the participants who reported having watched at least one
Khan academy course, 1 participant declined to report the number of courses
they had watched; 7 participants reported having watched 1--2 courses; 11
reported having watched 3--5 courses; 8 reported having watched 5--10 courses;
and 19 reported having watched 10 or more courses. We also asked participants
about the specific courses they had watched, categorized under different
subject areas. In the ``Mathematics'' area participants reported having watched
videos on AP Calculus AB (21 participants), Precalculus (17 participants),
Algebra 2 (14 participants), AP Calculus BC (12 participants), Trigonometry (11
participants), Algebra 1 (10 participants), Geometry (8 participants),
Pre-algebra (7 participants), Multivariable Calculus (5 participants),
Differential Equations (5 participants), Statistics and Probability (4
participants), AP Statistics (2 participants), Linear Algebra (2 participants),
Early Math (1 participant), Arithmetic (1 participant), and other videos not
listed in our survey (6 participants). In the ``Science and engineering'' area
participants reported having watched videos on Chemistry, AP Chemistry, or
Organic Chemstry (21 participants); Physics, AP Physics I, or AP Physics II (15
participants); Biology, AP Biology; or High school Biology (15 participants);
Health and Medicine (1 participant); or other videos not listed in our survey
(20 participants). We also asked participants if they had specifically seen the
videos used in our experiment. When we asked about the \textit{Four Fundamental
Forces} video, 45 participants reported not having watched it before, 1
participant reported that they were not sure if they had watched it before, and
4 participants declined to respond. When we asked about the \textit{Birth of
Stars} video, 46 participants reported not having watched it before and 4
participants declined to respond. When we asked participants about non-Khan
academy online courses, they reported having watched or taken courses on
Mathematics (15 participants), Science and engineering (11 participants), Test
preparation (9 participants), Economics and finance (3 participants), Arts and
humanities (2 participants), Computing (2 participants), and other categories
not listed in our survey (18 participants). Finally, we asked participants
about in-person courses they had taken in different subject areas. They
reported taking courses in Mathematics (39 participants), Science and
engineering (38 participants), Arts and humanities (35 participants), Test
preparation (27 participants), Economics and finance (26 participants),
Computing (15 participants), College and careers (7 participants), or other
courses not listed in our survey (6 participants).

@@

An important feature of our approach is that, given a trained text embedding
model and participants' quiz performance on each question, we can estimate
their knowledge about \textit{any} content expressable by the embedding model--
not solely the content explicitly probed by the quiz questions. To visualize
these estimates
(Figs.~\ref{fig:knowledge-maps},~\individualKnowledgeMapsA,~\individualKnowledgeMapsB,~\individualKnowledgeMapsC,~\individualLearningMapsA,
and~\individualLearningMapsB), we used UMAP~\citep{McInEtal18a} to define a 2D
projection of the text embedding space. Sampling the original 100-dimensional
space at high resolution to obtain an adequate set of topic vectors spanning
the embedding space would be computationally intractable. However, sampling a
2D grid is much more feasible. We defined a rectangle enclosing the 2D
projections of the lectures' and quizzes' embeddings, and we sampled points
from a regular $100 \times 100$ grid of coordinates that evenly tiled the
enclosing rectangle. We sought to estimate participants' knoweldge (and
learning---i.e., changes in knowledge) at each of the resulting 10000
coordinates.

@@

>>>>

\begin{abstract} We develop a mathematical framework, based on natural language
processing models, for tracking and characterizing the acquisition of
conceptual knowledge. Our approach embeds each concept in a high dimensional
representation space, where nearby coordinates reflect similar or related
concepts. We test our approach using behavioral data from participants who
answered small sets of multiple choice quiz questions interleaved between
watching two course videos from the Khan Academy platform. We applied our
framework to the videos' transcripts, and to text of the quiz questions, to
quantify the content of each moment of video and each quiz question. We used
these embeddings, along with participants' quiz responses, to track how the
learners' knowledge changed after watching each video. Our findings show how a
small set of quiz questions may be used to obtain rich and meaningful high
resolution insights into what each learner knows, and how their knowledge
changes over time as they learn.

@@

We asked our participants to answer questions from several multiple choice
quizzes and watch two lecture videos from the \textit{Khan Academy} platform
(Fig.~\ref{fig:experiment}). One lecture video, entitled \textit{Four
Fundamental Forces}, was about the four fundamental forces in physics: gravity,
strong and weak interactions, and electromagnetism. The second lecture video,
entitled \textit{Birth of Stars}, provides an overview of our current
understanding of how stars form. We selected both lessons to be (a) accessible
to a broad audience, e.g., by minimizing prerequisite knowledge, (b) largely
independent of each other, e.g., so that the two videos focused on different
material and did not depend on each other, and (c) related to each other, e.g.,
so that both videos contained at least \textit{some} similar or overlapping
content. The two videos we selected are introductory (i.e., minimizing specific
prerequisite knowledge), and are about different primary concepts, but they
both also touch on ``physics'' and ``astronomy'' themes. We also wrote a set of
multiple choice quiz questions that would enable us to test participants'
knowledge about each individual video and about related content not
specifically presented in either video (Tab.~\questions). Participants answered
questions randomly drawn from each content area (lecture 1, lecture 2, and
general physics knowledge) across each of three quizzes. Quiz 1 was intended to
assessed participants' knowledge before training; quiz 2 assessed knowledge
after watching the \textit{Four Fundamental Forces} video (i.e., lecture 1);
and quiz 3 assessed knowledge after watching the \textit{Birth of Stars} video
(i.e., lecture 2).

@@

Given that we trained the text embedding model using the video transcripts, we
wondered whether the questions that were (ostensibly, by design) ``about'' the
content of each lecture would ``match up'' correctly with the lectures. In
other words, we hoped that the text embeddings would capture something about
the deeper conceptual content of the lectures, beyond surface details such as
exact wording choices. If so, when we embed \textit{new} text outside of the
model's training set, we should see a correspondance between the embeddings of
the training data (i.e., snippets of text from the lectures' transcripts) and
other text that reflects related concepts (e.g., questions \textit{about} each
lecture). Further, although the content from any given moment from a lecture
might stray from the average content (across all timepoints), we hoped that
\textit{variability} in each topic's expression over timepoints within a
lecture would match up with the variability in topic expressions for questions
about that lecture. Intuitively, the variability in the expression of a given
topic relates to how much ``information''~\citep{Fish22} the lecture (or
questions) reflect about that topic. When we compared the variability in topic
weights across each lecture's timepoints with the variability in topic weights
across each question set, we found a strong correspondence
(Fig.~\ref{fig:topics}). The most variable topics from the \textit{Four
Fundamental Forces} lecture, and questions about that lecture, are 2, 3, 8, 12,
13, 14, and 15. The most variable topics from the \textit{Birth of Stars}
lecture, and questions about that lecture, are 1, 4, 5, 6, and 7. This strong
overlap between the lectures and questions specifically about each lecture
indicates that the topic model captures some of the underlying conceptual
content.

@@

Although a single lecture may be organized around a single broad theme at a
coarse scale, at a finer scale each moment of a lecture typically covers a
narrower range of content. We wondered whether a text embedding model trained
on the lectures' transcripts might capture some of this finer scale content.
For example, if a particular question asks about the content from one small
part of a lecture, we wondered whether our text embedding model could be used
to automatically identify the ``matching'' moment(s) in the lecture. When we
correlated each question's topic vector with the topic vectors for each second
of the lectures, we found some evidence that each question is temporally
specific (Fig.~\ref{fig:question-correlations}). In particular, most questions'
topic vectors were maximally correlated with a well-defined (and relatively
narrow) range of timepoints from their corresponding lectures, and the
correlations fell off sharply outside of that range. We also examined the
best-matching intervals for each question qualitatively by comparing the text
of the question to the text of the most-correlated parts of the lectures.
Despite that the questions were excluded from the text embedding model's
training set, in general we found (through manual inspection) a close
correspondence between the conceptual content that each question covered and
the content covered by the best-matching moments of the lectures. Two
representative examples are shown at the bottom of
Figure~\ref{fig:question-correlations}.

The ability to quantify how much each question is ``asking about'' the content
from each moment of the lectures could enable high-resolution insights into
participants' knowledge. Traditional approaches to estimating how much a
student ``knows'' about the content of a given lecture entail computing the
proportion of correctly answered questions. But if two students receive
identical scores on an exam, might our modeling framework help us to gain more
nuanced insights into the \textit{specific} content that each student has
mastered (or failed to master)? For example, a student who misses three
questions that were all about the same concept (e.g., concept $A$) will have
gotten the same \textit{proportion} of questions correct as another student who
missed three questions about three \textit{different} concepts (e.g., $A$, $B$,
and $C$). But if we wanted to fill in the ``gaps'' in the two students'
understandings, we might do well to focus on concept $A$ for the first student,
but to also add in materials pertaining to concepts $B$ and $C$ for the second
student.

@@

If we are able to accurately estimate a participant's knowledge about the
content tested by a given question, the estimated knowledge should have some
predictive information about whether the participant is likely to answer the
question correctly or incorrectly. For each question in turn, for each
participant, we used Equation~\ref{eqn:prop} to estimate (using all
\textit{other} questions from the same quiz, from the same participant) the
participant's knowledge at the held-out question's embedding coordinate. For
each quiz, we grouped these estimates into two distributions: one for the
estimated knowledge at the coordinates of each \textit{correctly} answered
question, and another for the estimated knowledge at the coordinates of each
\textit{incorrectly} answered question (Fig.~\ref{fig:predictions}). We then
used independent samples $t$-tests to compare the means of these distributions
of estimated knowledge.

@@

    \caption{\textbf{Mapping out the geometry of knowledge and learning.}
    \textbf{A. Average ``knowledge maps'' estimated using each quiz.} Each map
    displays a 2D projection of the estimated knowledge about the content
    reflected by \textit{all} regions of topic space (see \textit{Creating
    knowledge and learning map visualizations}). The topic trajectories of each
    lecture and the coordinates of each question are indicated by dotted lines
    and dots. Each map reflects an average across all participants. For
    individual participants' maps, see
    Figures~\individualKnowledgeMapsA,~\individualKnowledgeMapsB,
    and~\individualKnowledgeMapsC. \textbf{B. Average ``learning maps''
    estimated between each successive pair of quizzes.} The learning maps are
    in the same general format as the knowledge maps in Panel A, but each
    coordinate in the learning maps indicates the \textit{difference} between
    the corresponding coordinates in the indicated \textit{pair} of knowledge
    maps---i.e., how much the estimated knowledge ``changed'' across the two
    quizzes. Each map reflects an average across all participants. For
    individual participants' maps, see
    Figures~\individualLearningMapsA~and~\individualLearningMapsB. \textbf{C.
    Word clouds for sampled points in topic space.} Each word cloud displays
    the relative weights of each word reflected by the blend of topics
    represented at the locations of the stars in the maps. The words' colors
    indicate how much each word is weighted on average across all timepoints'
    topic vectors in the \textit{Four Fundamental Forces} (blue) and
    \textit{Birth of Stars} (green) videos, respectively.}

@@

Because the 2D projection we used to construct the knowledge and learning maps
is (partially) invertible, we may gain additional insights into the estimates
by reconstructing the original high-dimensional topic vectors for any point(s)
in the maps we are interested in. For example, this could serve as a useful
tool for an instructor looking to better understand which content areas a
student (or a group of students) knows well (or poorly). As a demonstration, we
show the top-weighted words from the blends of topics reconstructed from three
example locations on the maps (Fig.~\ref{fig:knowledge-maps}C): one point near
the \textit{Four Fundamental Forces} embedding (yellow); a second point near
the \textit{Birth of Stars} embedding (orange), and a third point somewhere in
between the two lectures' embeddings (pink). As shown in the word clouds in the
Panel, the top-weighted words at the example coordinate near the \textit{Four
Fundamental Forces} embedding also tended to be weighted heavily by the topics
expressed in that lecture. Similarly, the top-weighted words at the example
coordinate near the \textit{Birth of Stars} embedding tended to be weighted
most heavily by the topics expressed in \textit{that} lecture. And the
top-weighted words at the example coordinate between the two lectures'
embeddings show a roughly even mix of words most strongly associated with each
lecture.

@@

The notion of empathy also related to ``theory of mind'' of other
individuals~\citep{GoldWinn12, KansEtal15, Melt11}. Considering others' unique
perspectives, prior experiences, knowledge, goals, etc., can help us to more
effectively interact and communicate~\citep{ShaoEtal18, StepBaer06, Ratk18}.
The knowledge and learning maps we estimate in our study
(Fig.~\ref{fig:knowledge-maps}) hint at one potential form that an automated
``empathetic'' teacher might take. We imagine automated content delivery
systems that adapt lessons on the fly according to continually updated
estimates of what students know and how quickly they are learning different
conceptual content~\citep[e.g., building on ideas such as][and
others]{AndeSkwa86, Half88, Kuma05, WolzEtal88}.

@@

Participants' ages ranged from 18 to 22 years (mean: 19.52 years;
standard deviation: 1.09 years). A total of 15 participants reported
their gender as male and 35 participants reported their gender as
female. A total of 49 participants reported their native language as
``English'' and 1 reported having another native language. A total of
47 participants reported their ethnicity as ``Not Hispanic or Latino''
and three reported their ethnicity as ``Hispanic or Latino.''
Participants reported their races as White (32 participants), Asian
(14 participants), Black or African American (5 participants),
American Indian or Alaska Native (1 participant), and Native Hawaiian or
Other Pacific Islander (1 participant). (Note that some participants
selected multiple racial categories.)

@@

Participants reported their undergraduate major(s) as Social Sciences (28
participants), Natural sciences (16), Professional (e.g., pre-med or pre-law; 8
participants), Mathematics and engineering (7 participants), Humanities (4
participants), or Undecided (3 participants). Note that some participants
selected multiple categories for their undergraduate major. We also asked
participants about the courses they had taken. In total, 46 participants
reported having taken at least one Khan academy course in the past or being
familiar with the Khan academy, and 4 reported not having taken any Khan
academy courses. Of the participants who reported having watched at least one
Khan academy course, 1 participant declined to report the number of courses
they had watched; 7 participants reported having watched 1--2 courses; 11
reported having watched 3--5 courses; 8 reported having watched 5--10 courses;
and 19 reported having watched 10 or more courses. We also asked participants
about the specific courses they had watched, categorized under different
subject areas. In the ``Mathematics'' area participants reported having watched
videos on AP Calculus AB (21 participants), Precalculus (17 participants),
Algebra 2 (14 participants), AP Calculus BC (12 participants), Trigonometry (11
participants), Algebra 1 (10 participants), Geometry (8 participants),
Pre-algebra (7 participants), Multivariable Calculus (5 participants),
Differential Equations (5 participants), Statistics and Probability (4
participants), AP Statistics (2 participants), Linear Algebra (2 participants),
Early Math (1 participant), Arithmetic (1 participant), and other videos not
listed in our survey (6 participants). In the ``Science and engineering'' area
participants reported having watched videos on Chemistry, AP Chemistry, or
Organic Chemistry (21 participants); Physics, AP Physics I, or AP Physics II (15
participants); Biology, AP Biology; or High school Biology (15 participants);
Health and Medicine (1 participant); or other videos not listed in our survey
(20 participants). We also asked participants if they had specifically seen the
videos used in our experiment. When we asked about the \textit{Four Fundamental
Forces} video, 45 participants reported not having watched it before, 1
participant reported that they were not sure if they had watched it before, and
4 participants declined to respond. When we asked about the \textit{Birth of
Stars} video, 46 participants reported not having watched it before and 4
participants declined to respond. When we asked participants about non-Khan
academy online courses, they reported having watched or taken courses on
Mathematics (15 participants), Science and engineering (11 participants), Test
preparation (9 participants), Economics and finance (3 participants), Arts and
humanities (2 participants), Computing (2 participants), and other categories
not listed in our survey (18 participants). Finally, we asked participants
about in-person courses they had taken in different subject areas. They
reported taking courses in Mathematics (39 participants), Science and
engineering (38 participants), Arts and humanities (35 participants), Test
preparation (27 participants), Economics and finance (26 participants),
Computing (15 participants), College and careers (7 participants), or other
courses not listed in our survey (6 participants).

@@

An important feature of our approach is that, given a trained text embedding
model and participants' quiz performance on each question, we can estimate
their knowledge about \textit{any} content expressible by the embedding model--
not solely the content explicitly probed by the quiz questions. To visualize
these estimates
(Figs.~\ref{fig:knowledge-maps},~\individualKnowledgeMapsA,~\individualKnowledgeMapsB,~\individualKnowledgeMapsC,~\individualLearningMapsA,
and~\individualLearningMapsB), we used UMAP~\citep{McInEtal18a} to define a 2D
projection of the text embedding space. Sampling the original 100-dimensional
space at high resolution to obtain an adequate set of topic vectors spanning
the embedding space would be computationally intractable. However, sampling a
2D grid is much more feasible. We defined a rectangle enclosing the 2D
projections of the lectures' and quizzes' embeddings, and we sampled points
from a regular $100 \times 100$ grid of coordinates that evenly tiled the
enclosing rectangle. We sought to estimate participants' knowledge (and
learning---i.e., changes in knowledge) at each of the resulting 10000
coordinates.

@@

\section*{Author contributions}

Conceptualization: PCF, ACH, and JRM. Methodology: PCF, ACH, and JRM. Software:
PCF. Validation: PCF. Formal analysis: PCF. Resources: PCF, ACH, and JRM. Data
curation: PCF. Writing (original draft): JRM. Writing (review and editing):
PCF, ACH, and JRM. Visualization: PCH and JRM. Supervision: JRM. Project
administration: PCH. Funding acquisition: JRM.

\section*{Data and code availability}

All of the data analyzed in this manuscript, along with all of the code for
running our experiment and carrying out the analyses may be found at
https://github.com/Con\-text\-Lab/eff\-ic\-ient-learn\-ing-khan.

\section*{Acknowledgements}

We acknowledge useful discussions, assistance in setting up an earlier
(unpublished) version of this study, and assistance with some of the data
collection efforts from Will Baxley, Max Bluestone, Daniel Carstensen, Kunal
Jha, Caroline Lee, Lucy Owen, Xinming Xu, and Kirsten Ziman. Our work was
supported in part by NSF CAREER Award Number 2145172 to JRM. The content is
solely the responsibility of the authors and does not necessarily represent the
official views of our supporting organizations. The funders had no role in
study design, data collection and analysis, decision to publish, or preparation
of the manuscript.
